---
title: "The Relationship Between Secondary Education and Nobel Laureates Across Countries"
author: "Marlon Trejo"
date: today
date-format: long
abstract: "We consider whether a country's percentage of the population with at least some secondary schooling has any association on the number of nobel laureates that country's had. The regression line shows a slight upward trend, indicating that, on average, there is a positive correlation. Given the p-value, we can say we have enough evidence at the 5% significance level that there is a linear association between these two variables."
format: pdf
number-sections: true
bibliography: references.bib


references:
  - id: cdc2026
    type: webpage
    author:
      - literal: Centers for Disease Control and Prevention
    title: "Fatal Injury Reports, National and Regional, 2024"
    container-title: "WISQARS"
    issued:
      - year: 2026
    URL: "https://wisqars.cdc.gov/"

  - id: R_language
    type: software
    author:
      - literal: R Core Team
    title: "R: A Language and Environment for Statistical Computing"
    publisher: "R Foundation for Statistical Computing"
    publisher-place: "Vienna, Austria"
    URL: "https://www.R-project.org/"
---

```{r}
#| include: false
#| warning: false
#| message: false

data <- data.frame(
  countries = c(
    "United States",
    "United Kingdom",
    "Germany",
    "France",
    "Sweden",
    "Japan",
    "Russia/USSR",
    "Canada",
    "Switzerland",
    "Austria",
    "Netherlands",
    "Italy",
    "Poland",
    "Hungary",
    "Australia",
    "Israel",
    "Denmark",
    "Norway",
    "India",
    "Belgium",
    "Ireland",
    "South Africa",
    "China",
    "Spain",
    "Belarus",
    "Czechia",
    "Ukraine",
    "Argentina",
    "Finland",
    "Egypt",
    "Romania",
    "Taiwan",
    "Lithuania",
    "Mexico",
    "New Zealand",
    "South Korea",
    "Tunisia",
    "Turkey",
    "Algeria",
    "Azerbaijan",
    "Chile",
    "Colombia",
    "Cyprus",
    "East Timor",
    "Iran",
    "Greece",
    "Guatemala",
    "Liberia",
    "Luxembourg",
    "Pakistan",
    "Portugal",
    "Saint Lucia",
    "Venezuela",
    "Bangladesh",
    "Brazil",
    "Bulgaria",
    "Costa Rica",
    "Democratic Republic of the Congo",
    "Ethiopia",
    "Faroe Islands",
    "Ghana",
    "Hong Kong",
    "Iceland",
    "Iraq",
    "Jordan",
    "Kenya",
    "Latvia",
    "Lebanon",
    "North Macedonia",
    "Morocco",
    "Myanmar",
    "Nigeria",
    "Palestine",
    "Peru",
    "Philippines",
    "Saudi Arabia",
    "Tanzania",
    "Tibet",
    "Trinidad and Tobago",
    "Vietnam",
    "Yemen",
    "Yugoslavia",
    "Zimbabwe"
  ), 
  secondary_education = c(
    95.3,  # United States
    82.9,  # United Kingdom
    96.7,  # Germany
    82.5,  # France
    88.0,  # Sweden
    91.8,  # Japan
    94.6,  # Russia/USSR
    100.0, # Canada
    96.7,  # Switzerland
    98.9,  # Austria
    88.2,  # Netherlands
    82.3,  # Italy
    83.9,  # Poland
    96.6,  # Hungary
    91.5,  # Australia
    88.8,  # Israel
    89.5,  # Denmark
    95.3,  # Norway
    48.7,  # India
    82.3,  # Belgium
    85.5,  # Ireland
    74.9,  # South Africa
    75.0,  # China
    73.7,  # Spain
    89.3,  # Belarus
    99.8,  # Czechia
    95.1,  # Ukraine
    62.4,  # Argentina
    100.0, # Finland
    61.4,  # Egypt
    89.1,  # Romania
    NA,    # Taiwan
    93.1,  # Lithuania
    57.4,  # Mexico
    98.7,  # New Zealand
    91.4,  # South Korea
    43.9,  # Tunisia
    54.0,  # Turkey
    34.9,  # Algeria
    95.6,  # Azerbaijan
    76.5,  # Chile
    54.9,  # Colombia
    79.0,  # Cyprus
    NA,    # East Timor / Timor-Leste
    67.7,  # Iran
    67.5,  # Greece
    36.8,  # Guatemala
    28.3,  # Liberia
    100.0, # Luxembourg
    35.4,  # Pakistan
    51.4,  # Portugal
    45.2,  # Saint Lucia
    68.9,  # Venezuela
    43.1,  # Bangladesh
    57.5,  # Brazil
    94.2,  # Bulgaria
    54.2,  # Costa Rica
    24.6,  # Democratic Republic of the Congo
    15.8,  # Ethiopia
    NA,    # Faroe Islands
    59.8,  # Ghana
    77.4,  # Hong Kong
    98.6,  # Iceland
    45.6,  # Iraq
    81.3,  # Jordan
    32.0,  # Kenya
    99.1,  # Latvia
    54.2,  # Lebanon
    47.8,  # North Macedonia
    29.4,  # Morocco
    23.8,  # Myanmar
    NA,    # Nigeria
    58.8,  # Palestine
    61.5,  # Peru
    71.6,  # Philippines
    66.5,  # Saudi Arabia
    12.6,  # Tanzania
    NA,    # Tibet
    69.8,  # Trinidad and Tobago
    71.7,  # Vietnam
    24.4,  # Yemen
    NA,    # Yugoslavia
    57.7   # Zimbabwe
  
  ),
  nobel_prizes = c(
  428,  # United States
  145,  # United Kingdom
  116,  # Germany
  79,   # France
  34,   # Sweden
  33,   # Japan
  30,   # Russia/USSR
  29,   # Canada
  27,   # Switzerland
  25,   # Austria
  22,   # Netherlands
  21,   # Italy
  19,   # Poland
  16,   # Hungary
  15,   # Australia
  14,   # Israel
  14,   # Denmark
  14,   # Norway
  13,   # India
  11,   # Belgium
  11,   # Ireland
  11,   # South Africa
  8,    # China
  8,    # Spain
  6,    # Belarus
  6,    # Czechia
  6,    # Ukraine
  5,    # Argentina
  5,    # Finland
  5,    # Egypt
  4,    # Romania
  NA,   # Taiwan (Wikipedia gives a range of 1–4)
  3,    # Lithuania
  3,    # Mexico
  3,    # New Zealand
  3,    # South Korea
  3,    # Tunisia
  3,    # Turkey
  2,    # Algeria
  2,    # Azerbaijan
  2,    # Chile
  2,    # Colombia
  2,    # Cyprus
  2,    # East Timor
  2,    # Iran
  2,    # Greece
  2,    # Guatemala
  2,    # Liberia
  2,    # Luxembourg
  2,    # Pakistan
  2,    # Portugal
  2,    # Saint Lucia
  2,    # Venezuela
  1,    # Bangladesh
  1,    # Brazil
  1,    # Bulgaria
  1,    # Costa Rica
  1,    # Democratic Republic of the Congo
  1,    # Ethiopia
  1,    # Faroe Islands
  1,    # Ghana
  1,    # Hong Kong
  1,    # Iceland
  1,    # Iraq
  1,    # Jordan
  1,    # Kenya
  1,    # Latvia
  1,    # Lebanon
  1,    # North Macedonia
  1,    # Morocco
  1,    # Myanmar
  1,    # Nigeria
  1,    # Palestine
  1,    # Peru
  1,    # Philippines
  1,    # Saudi Arabia
  1,    # Tanzania
  1,    # Tibet
  1,    # Trinidad and Tobago
  1,    # Vietnam
  1,    # Yemen
  1,    # Yugoslavia
  1     # Zimbabwe
)
)


data_clean <- na.omit(data)

lm_fit <- lm(nobel_prizes ~ secondary_education, data = data_clean)

summary(lm_fit)
plot(
  data_clean$secondary_education,
  data_clean$nobel_prizes,
  xlab = "Percent of Population With at Least Some Secondary Education",
  ylab = "Nobel Prizes"
)

abline(lm_fit)

plot(data_clean$secondary_education,
     residuals(lm_fit),
     xlab = "Percent of Population With at Least Some Secondary Education",
     ylab = "Residuals",
     main = "Residuals vs Percentage",
     pch = 19)

abline(h = 0)

qqnorm(residuals(lm_fit))
qqline(residuals(lm_fit))

```

# Introduction

Nobel Prizes are awarded in accordance with the principle of "for the greatest benefit to humankind". The six categories are physics, chemistry, medicine, literature, peace, and economic science. This analysis was done to determine whether there was a linear relationship between the percentage of the population with at least some secondary education in that country, and the number of Nobel laureates in that country. 

The primary motivation is to inspire further research into what aspects of a country's education system causes high scientific/civil achievement. This paper is meant to see if education level and Nobel prizes won per country are correlated. If they aren't, that would indicate that there's other elements involved (such as talent or individual strengths) that produces Nobel laureates.

A regression line was fitted, and a slope of .529 was found. The p-value of .0328 indicates that, using a 5% significance level and a two-tailed test, we have enough evidence to conclude that the slope is different from 0. This is important, because it shows that there is a linear relationship between the two variables.
 

@sec-data introduces the data used in this analysis. @sec-methods describes the statistical model. @sec-results presents the results of the analysis, and @sec-discussion discusses the findings.

# Data {#sec-data}

The predictor variable data comes from the official Human Development Report from the United Nations from 2016 [@undp2016]. The outcome variable comes from the official Wikipedia list of Nobel Laureates per Country [@wikipedia_nobel_country]. The outcome variable is the number of Nobel Prize winners from a certain country. The predictor variable is a percentage, and it corresponds to the population in that country with at least some secondary education. This comes from surveys given to people aged 25 or older. Since there were some countries in the Nobel Prize data but missing in the data set, those countries were erased, instead of getting the data from an additional data source (to keep variables consistent). There are obvious limitations in using surveys-people could lie or data could be incomplete, 

# Methods {#sec-methods}

I adopted the simple linear regression model

$$
Y_i = \beta_0 + \beta_1 X_i + \varepsilon_i,
$$

where $Y_i$ represents the response variable (Number of Nobel Prize Winners), and $X_i$ represents the predictor variable (Percent of Population with at Least Some Secondary Education).

In this model, $\beta_0$ doesn't mean anything, since it represents the expected number of Nobel prize winners from a country with no percent of the population having some secondary education. $\beta_1$ represents the expected change in the number of Nobel Prize winners associated with an increase of one percentage point of the population having some secondary education. $\varepsilon_i$ represents elements that cause the count to differ from the expected value (errors).

I implemented this analysis using the R programming language [@R_language]. Specifically, I used the lm() function in R.

The assumptions of this model include normally distributed and errors, which we will show to be true later. Potential weaknesses of this approach might be that the variance doesn't appear to be constant-the higher the percentage of individuals with at least some secondary education, the more spread out the number of Nobel Prize winners. 

 
# Results {#sec-results}

The estimated regression model is

$$
\widehat{Y}
=
`r round(coef(lm_fit)[1], 3)`
+
`r round(coef(lm_fit)[2], 3)`X.
$$

The estimated slope parameter is $b_1 =$ `r round(coef(lm_fit)[2], 3)`.

In other words, for each additional increase in the percent of the population with at least some secondary education, the number of Nobel prize winners in that country changes by approximately `r round(coef(lm_fit)[2], 3)` winners, on average.

The estimated intercept is `r round(coef(lm_fit)[1], 3)`.

We will do no confidence intervals or hypothesis testing, since they would not be reliable due to the non-constant variance shown by the residuals. 

@fig-suicide-residual shows the residuals from the fitted regression model. The points seem to have a funnel shape near the right, indicating non equal variance. The data does seem to be normally distributed, as shown in Figure 3 below.



```{r}
#| label: fig-suicide-residual
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Residuals from the simple linear regression model plotted"

plot(data_clean$secondary_education,
     residuals(lm_fit),
     xlab = "Percent of Population With at Least Some Secondary Education",
     ylab = "Residuals",
     main = "Residuals vs Percentage",
     pch = 19)

abline(h = 0)
```

```{r}
#| label: fig-suicide-scatter
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Percent of population with some secondary education vs. Nobel Prize winners in that country with fitted linear regression line."

par(mar = c(8, 4, 4, 2))

plot(
  data_clean$secondary_education,
  data_clean$nobel_prizes,
  xlab = "Percent of Population With at Least Some Secondary Education",
  ylab = "Nobel Prizes"
)
abline(lm_fit)
```
```{r}
#| label: Q
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "QQ Norm Plot."

par(mar = c(8, 4, 4, 2))


qqnorm(residuals(lm_fit))
qqline(residuals(lm_fit))
```
 

# Discussion {#sec-discussion}
This paper attempted to see if there was a linear relationship between Nobel prize winners per country and the percentage of people with at least some secondary education in that country. Based on our methods used, there seems to be a positive linear relationship. There is a slight upward slope, but a p-test can't be done due to non-constant variance of the errors. There are some weaknesses in this study, including the variance of the errors and the fact that the surveys done probably aren't 100% accurate.Future work might be better handled by a more accurate predictor variable and more data.
 
 

 